---
title: "Day 4 notes - functions"
author: "Erica Bishop"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Nested for loop examples

```{r}

file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}

##paste automatically includes a space, paste0 concatenates without spaces
# inner for loop is evaluated first fully (i = 1, and j evaluates fully through to get i1,j1; i1,j2; i1,j3...etc. before moving on to i=2, j1, i2,j3... etc.)

#number f

```

Another example:

At each

```{r}

odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for (i in seq_along(odds)) {
  for (j in seq_along(evens)) {
    print(odds[i] * evens[j])
  }
}


```

# Functions

functions are self-contained bits of code that accept user input(s), do something with it, and return a useful output

consider using a function when you find yourself copying and pasting a bit of code more than twice

**Benefits of writing functions**

1.  you can give a function a usefule and readable name

2.  As requirements change, you only need to update code in one place instead of many

3.  You eliminated the chance of making incidental mistakes when you copy and paste

Minimizing lines of code makes it easier to read and troubleshoot

Functions will look like:

function_name \<- function (bird, dog) { pets = bird + dog return(pets) }

start by naming function, then specify arguments within the function() function, then specify what to do with arguments within {}

In python, use def function_name to name function, followed by arguments (argument 1, argument 2): what to do,

```{r}
#Example

exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 25)

exclaim_age("four")

##this function will take numeric or string, QUESTION: how would I specify that I want my function to only take one class of data

```

**Naming functions**

-   Name of the function is important, it's better to be clear than concise when you have to make the trade-off

-   generally, function names should be verbs, and arguments should be nouns

    -   however, a noun could be the right choice when accessing properties, such as "mean" "coefficient," , etc.

    -   if you're using nouns like "get", "compute", or "calculate" then a noun could be better

-   function name should make it clear what it's doing

## Function Practice

