---
title: "Day 4 notes - functions"
author: "Erica Bishop"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Nested for loop examples

```{r}

file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}

##paste automatically includes a space, paste0 concatenates without spaces
# inner for loop is evaluated first fully (i = 1, and j evaluates fully through to get i1,j1; i1,j2; i1,j3...etc. before moving on to i=2, j1, i2,j3... etc.)

#number f

```

Another example:

At each

```{r}

odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for (i in seq_along(odds)) {
  for (j in seq_along(evens)) {
    print(odds[i] * evens[j])
  }
}


```

# Functions

functions are self-contained bits of code that accept user input(s), do something with it, and return a useful output

consider using a function when you find yourself copying and pasting a bit of code more than twice

**Benefits of writing functions**

1.  you can give a function a usefule and readable name

2.  As requirements change, you only need to update code in one place instead of many

3.  You eliminated the chance of making incidental mistakes when you copy and paste

Minimizing lines of code makes it easier to read and troubleshoot

Functions will look like:

function_name \<- function (bird, dog) { pets = bird + dog return(pets) }

start by naming function, then specify arguments within the function() function, then specify what to do with arguments within {}

In python, use def function_name to name function, followed by arguments (argument 1, argument 2): what to do,

```{r}
#Example

exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 25)

exclaim_age("four")

##this function will take numeric or string, QUESTION: how would I specify that I want my function to only take one class of data

```

**Naming functions**

-   Name of the function is important, it's better to be clear than concise when you have to make the trade-off

-   generally, function names should be verbs, and arguments should be nouns

    -   however, a noun could be the right choice when accessing properties, such as "mean" "coefficient," , etc.

    -   if you're using nouns like "get", "compute", or "calculate" then a noun could be better

-   function name should make it clear what it's doing

## Function Practice

Example: you run two miles and these are your quarter mile splits in minutes - write a for loop that returns total time for any two consecutive quarter mile splits

```{r}

quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)) {
  half_split <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_split)
}


```

**Functions with conditions**

can create a function with conditional statements

*Example: create a function that doubles the input*

```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}
```

*now use the function*

```{r}
double_this(200)

double_this("value")
##error with non-numeric argument to binary operator

#but can use on a vector:

double_this(c(1, 2, 3, 4, 5))

```

*Example: write a function that asks for the users name and age and returns an exclamation

```{r}

age_name_exclaim <- function(name, age) {
  print(paste("My name is", name, "and I'm", age, "years old!"))
}

age_name_exclaim("Erica", 25)

##don't have to spell out arguments, but helps to be explicit when reading code back:

age_name_exclaim(name = "Teddy", age = 4)

```

Another example: 

```{r}
calc_animal_age <- function(animal, age_hyrs) {
  if (animal == "dog") {
    print(age_hyrs * 7)
  } else if (animal == "goat") {
    print(age_hyrs * 4.7)
  } else { print("Error: only converts dog or goat ages")}
}

calc_animal_age("Dog", 2)

calc_animal_age("Goat", 3)

calc_animal_age("horse", 21)
##silent error with invalid argument

```


Another example:

```{r}

dog_name <- c("Teddy", "Waffle", "Banjo")
dog_food <- c("salmon", "pancakes", "chicken")

dog_choice <- tribble(
  ~dog_name, ~dog_food,
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

##now make a function:

dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$dog_food))
}

dog_menu(name = "Waffle")
#now this function will return data frame with only the row called by dog name

```















